---
title: "Descriptive stats for constructs and elements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Descriptive stats for constructs and elements}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(collapse = TRUE, comment = "#")
options(width=120)
library(OpenRepGrid)
```

Sometimes basic descriptive measures for constructs and elements, e.g. mean, standard deviation, are neeeded. To prompt descriptive statistics for the constructs and elements of a grid use the function `statsConstructs` and `statsElements`. The following measures are returned:

* item name
* item number
* number of valid cases
* mean
* standard deviation
* trimmed mean (with trim defaulting to `.1`)
* median (standard or interpolated)
* mad: median absolute deviation (from the median)
* minimum
* maximum
* skew
* kurtosis
* standard error


# R-Code

The following examples are identical for `statsElements`. Just replace `statsConstructs` by `statsElements` in order to analyze elements.

```{r}
d <- statsConstructs(fbb2003)
d
```

The returned object is a dataframe, so you may access them as usual. E.g. to retrieve the means of the constructs, type

```{r}
d$mean
```


```{r}
statsConstructs(fbb2003, trim = 10)
```

```{r}
statsConstructs(fbb2003, index=F)
```