---
title: "Elements correlations"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
csl: apa.csl
bibliography: literature.bib
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Elements correlations}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#")
library(OpenRepGrid)
options(width=120)
settings(show.scale=FALSE, show.meta=FALSE)
```

### Description

As a measure for element similarity correlations between elements are frequently used. Note that product moment correlations as a measure of similarity are flawed as they are not invariant to construct reflection
[@mackay_identification_1992; @bell_note_2010]. A correlation index invariant to construct reflection is Cohen's rc measure [-@cohen_rc:_1969], which can be calculated using the argument `rc=TRUE` which is the default option.

### R-Code

As a default the construct reflection invariant correlation Cohen's rc is calculated. 
                                       
Note how the values change when the simple product-moment correlation is used instaed of Cohen's rc. Make sure you know what you are doing when setting `rc=FALSE`.

```{r}
elementCor(mackay1992, rc=F)
```

Although nor recommended different measures, not invariant to construct relfection, can be prompoted, when setting `rc=FALSE`. 

* **Pearson** or Product-moment correlation (PMC) 
* **Kendall**'s tau rank correlation
* **Spearman**'s rank correlation

To request these types of correlations use the `method` argument plus `rc=FALSE`.

```{r results='hide'}
elementCor(mackay1992, rc=FALSE, meth="kendall")    # Kendalls tau correlation
elementCor(mackay1992, rc=FALSE, meth="spearman")   # Spearman rank correlation
```

Several arguments to format the output are available.

```{r}
elementCor(mackay1992, index=F, trim=6)
```

If the correlations are calculated for further processing, the correlations can be saved.

```{r}
r <- elementCor(mackay1992)
```

The object is matrix, so you can eassily acces the results. E.g. all correlations with the element `Self`.

```{r}
r[1, ]
```


### Literature

<!-- 
* Bell, R. C. (2010). A note on aligning constructs. Personal Construct Theory & Practice, (7), 42-48.
* Cohen, J. (1969). rc: A profile similarity coefficient invariant over variable reflection. Psychological Bulletin, 71(4), 281-284.
* Mackay, N. (1992). Identification, Reflection, and Correlation: Problems in the bases of Repertory Grid measures. International Journal of Personal Construct Psychology, 5(1), 57-75.
-->
