---
title: "Conflict measures"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 5
csl: apa.csl
bibliography: literature.bib
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Conflict measures}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#")
options(width = 120)
library(OpenRepGrid)
settings(show.scale = FALSE, show.meta = FALSE)
```


### Conflict measures bases on correlations

#### Description

The first approach to mathematically derive a conflict measure based on grid data was presented by [-@slade_measurement_1979]. They based their operationalization on an approach by @lauterbach_assessing_1975 who applied Heider's [-@heider_attitudes_1946] balance theory for a quantitative assessment of psychological conflict. The measure uses inter-construct correlation as a basis. It assess via the sign of the correlations of a triad of constructs if the constructs triad is balanced or imbalanced. A triad is imbalanced if one or three of the correlations are negative, i. e. when they leading to contrary implications [@slade_measurement_1979]. The function `indexConflict1` in `OpenRepGrid` calculates the proportion of conflictive triads in the grid as devised by Slade and Sheehan.

Slade and Sheehan’s approach has several drawbacks, as it does not take into account the magnitude of the inter-construct correlations. As a consequence significant and insignificant correlations are not differentiated. Also, correlations of small magnitude, i.e. near zero, which may have a positive or negative signe due to chance alone distort the measure [@winter_construct_1982; @bassler_new_1992]. An improved version of the index that overcomes these shortcomings was proposed by [-@bassler_new_1992] and incorporated into the program *CT* (correlation test). The index enhances the identification of imbalanced triads by using a criterion that incorporates the magnitudes of correlations. From a psychological viewpoint, it remains unclear though what is measured. Connections to the concept of cognitive differentiation have been presumed [@krauthauser_new_1994]. The function `indexConflict2` in `OpenRepGrid` calculates the proportion of conflictive triads in the grid as devised by [-@bassler_new_1992].


#### Calculation 

##### @slade_measurement_1979 approach 

Multiply all correlations from a triad (formed by the constructs A, B and C) $r_{AB} r_{AC} r_{BC}$. If the sign of the result is negative, the triad is imbalanced. 

The table below shows when a triad made up of the constructs A, B, and C is balanced and imbalanced.
 
$r_{AB}$  | $r_{AC}$ | $r_{BC}$ | Triad characteristic
--------- | --------- | ---- | -----
+ | + | + | balanced
+ | + | - | imbalanced
+ | - | + | imbalanced
+ | - | - | balanced
- | + | + | imbalanced
- | + | - | balanced
- | - | + | balanced
- | - | - | imbalanced


##### @bassler_new_1992 approach 

Order correlations of the triad by absolute magnitude, so that $r_{max} > r_{mdn} > r_{min}$.

Apply Fisher's Z-transformation and devision by 3 to yield values between 1 and -1
Check whether the triad is balanced by assessing if the following relation holds:

* If $Z_{max} Z_{mdn} > 0$, the triad is balanced if $Z_{max} Z_{mdn} - Z_{min} <= crit$
* If $Z_{max} Z_{mdn} < 0$, the triad is balanced if $Z_{min} - Z_{max} Z_{mdn} <= crit$

#### R-Code

To use the Slade and Sheehan approach use

```{r}
indexConflict1(boeker)
```

For the Bassler et al. approach use

```{r}
indexConflict2(boeker)
indexConflict2(boeker, crit = .05) # change critical value
```


### Conflict measures based on distances

#### Description

Measure of conflict or inconsistency as proposed by @bell_new_2004. The identification of conflict is based on distances rather than correlations as in other measures of conflict indexConflict1 and indexConflict2. It assesses if the distances between all components of a triad, made up of one element and two constructs, satisfies the "triangle inequality" [cf. @bell_new_2004]. If not, a triad is regarded as conflictive. An advantage of the measure is that it can be interpreted not only as a global measure for a grid but also on an element, construct, and element by construct level making it valuable for detailed feedback. Also, differences in conflict can be submitted to statistical testing procedures.

#### R-Code

```{r}
indexConflict3(leach2001a)
```

### Literature

<!-- LITERATURE VERBOSE BACKUP
Bassler, M., Krauthauser, H., & Hoffmann, S. O. (1992). A new approach to the identification of cognitive conflicts in the repertory grid: An illustrative case study. Journal of Constructivist Psychology, 5(1), 95-111.
Heider, F. (1958). The Psychology of Interpersonal Relation. John Wiley & Sons.
Krauthauser, H., Bassler, M., & Potratz, B. (1994). A New Approach to the Identification of Cognitive Conflicts in the Repertory Grid: A Nomothetic Study. Journal of Constructivist Psychology, 7(4), 283-299.
Lauterbach, W. (1975). Assessing psychological conflict. The British Journal of Social and Clinical Psychology, 14(1), 43-47.
Slade, P. D., & Sheehan, M. J. (1979). The measurement of 'conflict' in repertory grids. The British Journal of Psychology, 70(4), 519-524.
Winter, D. A. (1982). Construct relationships, psychological disorder and therapeutic change. The British Journal of Medical Psychology, 55 (Pt 3), 257-269.
-->
