---
title: "Settings"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Settings}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(collapse = TRUE, comment = "#")
options(width = 120)
library(OpenRepGrid)
```

You can change several global settings for OpenRepGrid which will influence the way grids are displyed. You can promt the available settings by calling `settings()`.

```{r}
settings()
```

By default information about the scale and grid meta data is displayed.

```{r}
fbb2003
```

To change the settings enter the new parameters as `name=value` pairs. To turn off the display of the scale and meta information type in

```{r}
settings(show.scale = FALSE, show.meta = FALSE)
```

Now all grids are displayed without this information:

```{r}
fbb2003
```

Other settings that can be modified are maximal length of the construct that is displayed. Changing `show.cut`to `40` will prevent that the names of the constructs are cut off.

```{r}
settings(show.cut = 40)
fbb2003
```

In order to not print the number of constructs and elements you can turn off this options by 

```{r}
settings(c.no = FALSE, e.no = FALSE)
fbb2003
```
